\title{Reading Quiz 1 for Electromagnetic Theory (PHYS330)}
\author{Dr. Jordan Hanson - Whittier College Dept. of Physics and Astronomy}
\date{\today}
\documentclass[10pt]{article}
\usepackage[a4paper, total={18cm, 27cm}]{geometry}
\usepackage{outlines}
\usepackage{hyperref}
\begin{document}
\maketitle

\begin{abstract}
A summary of content covered in chapter 1 of Introduction to Electrodynamics. 
\end{abstract}
\noindent

\section{Vectors and Scalars}

\begin{enumerate}
\item Prove that scalar multiplication distributes over vector addition: $a(\vec{B} + \vec{C}) = a\vec{B} + a\vec{C}$.  \textit{Hint: break into components.} \\ \vspace{1cm}
\item What is wrong with this combination of objects?  $\nabla(f(x,y) + \nabla g(x,y))$ \\ \vspace{1cm}
\item (a) Create a vector field that has zero curl.  (b) Perform the line integral with the unit circle in the xy-plane as the path.  Does the result make sense? (c) What is the divergence of your field? \\ \vspace{2.5cm}
\end{enumerate}

\section{Vector Rotations}

\begin{enumerate}
\item Imagine a Cartesian coordinate system (right-handed) with $\hat{x}$ right, $\hat{y}$ up, and $\hat{z}$ out of the page. (a) Define a vector in the xy-plane.  (b) Rotate the coordinate system \textit{about the z-axis} by 45 degrees, counter-clockwise.  What are the new components of your vector?  (c) Determine if it has the same magnitude (it should). \\ \vspace{3cm}
\end{enumerate}

\section{Fundamental Theorems}

\begin{enumerate}
\item \textbf{Stokes' Theorem}.  Let $\vec{v} = s^{-1} \hat{\phi}$ in cylindrical coordinates, where the three coordinates in the text are $(s,\phi,z)$.  What is the result of $\int_{S} (\nabla \times \vec{v}) \cdot d\vec{a}$, if the surface is the \textit{closed sphere} with radius 1, centered on the origin.  \textit{Hint: This problem is a trick...What happens when Stokes' Theorem gets a closed surface?  Don't overthink it!}
\end{enumerate}

\clearpage

\section{Dirac Delta Functions}

\begin{enumerate}
\item Let
\begin{equation}
f(x) \star g(x) = \left(\frac{f(x) - g(x)}{f(x) + g(x)}\right)
\end{equation}
Evaluate $\int_{-\infty}^{\infty} (f(x) \star g(x)) \delta(x) dx$, if:
\begin{itemize}
\item $f(x) = \cos(x)$ and $g(x) = \sin(x)$.
\item $f(x) = \cosh(x)$ and $g(x) = \sinh(x)$.
\item $f(x) = a + ax + ax^2 + ...$, and $g(x) = b + bx + bx^2 + ...$
\end{itemize}
\end{enumerate}

\end{document}
